Introduction to Moltres: an Application for Simulation of Molten Salt Reactors

### 2017-11-21 Moltres.tex
english edits

\gls{MSR} modeling efforts describe steady-state and
transient behavior. Krepel et al. extended the in-house \gls{LWR}
diffusion code DYN3D to consider drift of delayed neutron precursors alongside
the reactor temperature profile, re-casting the extended code as
DYN3D-MSR \cite{krepel_dyn3d-msr_2007}. That work compared DYN3D-MSR against
experimental \gls{MSRE} data and then used it to simulate local fuel channel
blockages as well as local temperature perturbations.

In a similar vein, Kophazi et. al. used iterative coupling between in-house
three-dimensional neutronic and one-dimensional heat conduction models DALTON
and THERM to analyze normal \gls{MSRE} operation as well as
channel-blocking-incident transients \cite{kophazi_development_2009}. The
Kophazi model added entrance effects of heat transfer coefficients as well as
thermal coupling between fuel channels through moderator heat conduction. More
recently, Cammi et. al. performed a 2D-axisymmetric single-channel analysis of
the \gls{MSBR} using the commercial finite element package COMSOL Multiphysics
\cite{cammi_multi-physics_2011}. That work directly solved the fuel salt
velocity field, used heterogeneous group constants in fuel and moderator
regions, and employed the \gls{COMSOL} software package intrinsically designed
for coupled multi-physics simulation.  Fiorina, Lathouwers, and their
colleagues conducted a benchmarking exercise \cite{fiorina_modelling_2013} in
which this Politecnico di Milano approach was expanded to a multi-channel model
and compared to code from the University of Delft
\cite{de_zwaan_static_2007,van_der_linden_2012} based on the approach in
\cite{kophazi_development_2009}. These models showed good agreement for
accident transients in the \gls{MSFR}.

More recently, Aufiero et.  al \cite{aufiero_development_2014} have begun to
approach transient simulations in the \gls{MSFR} within the finite volume
OpenFOAM multiphysics toolkit\cite{weller_tensorial_1998}.  This approach
benefits from pre-implemented turbulence models availble in the OpenFOAM
library and captures the full-core three-dimensional geometry of the reactor
primary circuit.  OpenFOAM \gls{CFD} has additionally been shown by Laureau et
al. \cite{laureau_transient_2017} to couple well with Transient Fission Matrix
neutronics within the \gls{MSFR}.

The present work introduces the open source simulation tool, Moltres
\cite{lindsay_moltres_2017}, for simulating \glspl{MSR}.  By implementing
deterministic neutronics and thermal hydraulics in the context of the
\gls{MOOSE} finite element modeling framework, Moltres solves arbitrary-group
neutron diffusion, temperature, and precursor governing equations in anywhere
from one to three dimensions and can be deployed on an arbitrary number of
processing units.

@@

\subsection{Scaling study}

As described earlier, Moltres runs naturally in parallel. To test its
performance, we conducted a scaling study with a typical 2D axisymmetric input
file. \Cref{fig:scaling_study} shows Moltres' performance. The solution
time follows the ideal closely; slight deviations are expected due to
the expense of communication and the nature of iterative solvers. More
parallelism means the solution process uses more ``older'' information, 
which leads to slower convergence in terms of iterations 
\cite{satish_balay_petsc_2015}.

\begin{figure}[htpb]
  \centering
  \includegraphics{scaling_study.eps}
  \caption{Moltres performance on a typical 2D axisymmetric problem for
    different numbers of processors}
  \label{fig:scaling_study}
\end{figure}

>>>>

\gls{MSR} modeling efforts describe steady-state and
transient behavior. Krepel et al. extended the in-house \gls{LWR}
diffusion code DYN3D to consider drift of delayed neutron precursors alongside
the reactor temperature profile, re-casting the extended code as
DYN3D-MSR \cite{krepel_dyn3d-msr_2007}. That work compared DYN3D-MSR against
experimental \gls{MSRE} data and then used it to simulate local fuel channel
blockages as well as local temperature perturbations. In a similar vein, Kophazi
et. al. used iterative coupling between in-house three-dimensional neutronic and
one-dimensional heat conduction models DALTON and THERM to analyze normal 
\gls{MSRE} operation as well as channel-blocking-incident
transients \cite{kophazi_development_2009}. The Kophazi model added entrance 
effects of heat transfer coefficients as well as thermal
coupling between fuel channels through moderator heat conduction. More recently,
Cammi et. al. performed a 2D-axisymmetric single-channel analysis of the
\gls{MSBR} using the commercial finite element package COMSOL
Multiphysics \cite{cammi_multi-physics_2011}. That work directly solved the 
fuel salt velocity field, used heterogeneous group constants
in fuel and moderator regions, and employed a software package (COMSOL)
intrinsically designed for coupled multi-physics simulation.
Additionally, Aufiero et. al \cite{aufiero_development_2014} have begun to 
approach transient simulations in the \gls{MSFR} by directly coupling Serpent 2 
Monte Carlo neutronics with OpenFOAM \cite{weller_tensorial_1998} thermal-hydraulics.

The present work introduces the open source simulation tool, Moltres
\cite{lindsay_moltres_2017}, for simulating \glspl{MSR}.  By implementing
deterministic neutronics and thermal hydraulics in the context of the
\gls{MOOSE} finite element modeling framework, Moltres solves arbitrary-group
neutron diffusion, temperature, and precursor governing equations in anywhere
from one to three dimensions and can be deployed on an arbitrary number of
processing units. Through its computing power, Moltres is devoted to previously
unmatched fidelity in coupled neutronics and thermal hydraulics \gls{MSR}
simulation.

@@

\subsection{Scaling performance}

Parallelization in Moltres is implemented via LibMesh, which includes a set of
utilities for massively parallel finite element based computations, including
mesh input/output, a finite element library, and an interface for connections with
solver packages. Employing LibMesh provides Moltres
significant flexibility including the ability to swap out solver libraries
such as PetSc, which includes an expanding suite of parallel linear and
nonlinear solvers. Problem domain decomposition relies on LibMesh mesh
adaptation capabilities for running on a specific number of processors and
can either be performed manually before the start of the simulation or
automatically at the parameters of computation.

We conducted strong and weak scaling studies to characterize parallel 
performance in Moltres.
In case of strong scaling, the problem size remains fixed but the number of
processors is increased. Strong scaling studies seek to identify an
optimal ratio between the number of processors and elements for the most
rapid and power-efficient computation for a given problem. We measured
Moltres strong scaling with a simple 2D axisymmetric
case for various problem sizes separately for intra-node (2,820; 5,640;
11,280 and 28,200 elements) and extra-node (86,655; 173,310; 317,735;
664,355 elements) setup on Blue Waters' XK7 nodes (1 AMD 6276 Interlagos 
CPU, 16 integer scheduling units, 32 virtual units, nominal clock speed
2.3 GHz).

\Cref{fig:intra_strong_scaling} shows the simulation speed in seconds
per element vs. the number of cores on 1 node (maximum 32 cores). Up to 8 
cores, larger problems required considerably more time per
element because of cache overhead. However, beyond 8 cores, scaling demonstrates
asymptotic dependence on the number of processors due to increasing
communication costs. The best parallel efficiency for the intra-node study
is approximately 89\% and has been achieved for the largest problem
(28,200 elements).

\begin{figure}[htpb]
  \centering
  \includegraphics[width=\maxwidth{\textwidth}]{intra-node_strong.png}
  \caption{Moltres intra-node strong scaling efficiency for various problem 
        sizes, for $n_{cores} \in [1,32]$.}
  \label{fig:intra_strong_scaling}
\end{figure}

\Cref{fig:extra_strong_scaling} shows Moltres strong scaling up to 768
processors. This takes into account communication costs between nodes.
Similar to the intra-node study, when fewer than 128 cores were used, 
cache overhead causes performance slow down for larger problems. However, 
beyond 256 cores, the simulation time per element remains almost constant for small cases (86,655 and
173,310 elements) and slighly decreases for the two larger problems.
For extra-node scaling, parallel efficiency also grows with the
problem size and reaches an optimal value of 73\% for 664,355 elements.

\begin{figure}[htpb]
  \centering
  \includegraphics[width=\maxwidth{\textwidth}]{extra-node_strong.png}
  \caption{Moltres extra-node strong scaling efficiency for various problem size, for $n_{nodes} \in [1,24]$.}
  \label{fig:extra_strong_scaling}
\end{figure}

For the weak scaling study, the part of the problem (workload) assigned
to each processor stays constant and additional elements are used to
solve a bigger problem which would not fit in memory on a single node.
Thus, the weak scaling measurement is justification for memory-bound
application such as multiphysics code. Linear weak scaling is achieved
when the execution time stays constant while the workload increasing
in direct proportion to the number of cores. We performed Moltres
weak-scaling tests on Blue Waters, keeping the workload constant at
581, 985, 1970 and 3940 elements per core. \Cref{fig:intra_weak} shows
Moltres weak scaling performance measured for $n_{cores} \in [1,32]$ within
one Blue Waters node and \Cref{fig:extra_weak} demonstrates performance
for $n_{cores} \in [32,128]$. As expected, the largest drop in performance 
occurs when the number of cores increases from one to $\approx$ 8, which 
corresponds to switching from no communication to a 2-D domain
decomposition. The further reduction in performance of only about 50\%
over a range of 32 cores is likely caused by increased communication
latency appearing from collective \gls{MPI} calls. In the extra-node case,
the performance drops by a factor of three, which is most likely due to poor node
selection by the Blue Waters job scheduler and significantly increased latency and
bandwidth costs.

\begin{figure}[htpb]
  \centering
  \includegraphics[width=0.75\textwidth]{intra-node_weak.png}
  \caption{Weak scaling performance of Moltres on Blue Waters, in seconds per element vs. number of processors, for a constant number of elements
        per processor, and $n_{cores}\in[1,32]$.}
  \label{fig:intra_weak}
\end{figure}

\begin{figure}[htpb]
  \centering
  \includegraphics[width=0.75\textwidth]{extra-node_weak.png}
  \caption{Weak scaling performance of Moltres on Blue Waters, in seconds per element vs. number of processors, for a constant number of elements
        per processor, for and $n_{cores}\in[32,128]$.}
  \label{fig:extra_weak}
\end{figure}

Moltres scalability study results clearly indicate that parallelization
using LibMesh's automatic domain decomposition is good, but not perfectly
efficient. This scaling performance is satisfactory for \gls{MSR} simulations 
approached thus far and improved parallel performance would require further optimization
within LibMesh.  Moreover, Moltres is memory-bound and therefore very sensitive to host memory and memory
bandwidth. Consequently, if improved performance is needed, one could consider a transition from CPUs
computing to GPU-accelerated computing because GPUs operate on the fly with 
global memory, avoiding CPU cache storage issues. Another way
to improve parallel performance is to force the solver to use
``older'' information from previous iterations. However, this has been shown to 
slow convergence in terms of iterations and increased workload \cite{satish_balay_petsc_2015}.
